---
title: "Querying RIBBiTR DB for qPCR Swas Info"
author: "Brandon D. Hoenig"
date: "2024-04-22"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Load Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require(librarian)){
  install.packages('librarian')
  library(librarian)
}

shelf(tidyverse, 
      RPostgres, 
      DBI, 
      dbplyr)
library(janitor)
library(plotly)
library(htmlwidgets)
```

## Connect to the `ribbitr` database

```{r}
ribbitr_connection <-
  dbConnect(drv = dbDriver("Postgres"), 
            dbname = Sys.getenv('dbname'), 
            host = Sys.getenv('host'), 
            port = Sys.getenv('port'),
            user = Sys.getenv('user'),
            password = Sys.getenv('password'))

```

## Set `search_path` for `survey_data` schema
```{r}

search_path <-
  "set search_path = 'survey_data';"

dbExecute(conn = ribbitr_connection, 
          statement = search_path)
```


## Save each table needed 
```{r}
location_table <-
  tbl(src = ribbitr_connection, 
      from = "location") %>%
  collect() 

region_table <-
  tbl(src = ribbitr_connection, 
      from = "region") %>%
  collect()

site_table <-
  tbl(src = ribbitr_connection, 
      from = "site") %>%
  collect()
  
visit_table <-
  tbl(src = ribbitr_connection, 
      from = "visit") %>% 
  collect()
  
survey_table <-
  tbl(src = ribbitr_connection, 
      from = "survey") %>%
  collect()

capture_table <-
  tbl(src = ribbitr_connection, 
      from = "capture") %>%
  collect()

qpcr_bd_results_table <-
  tbl(src = ribbitr_connection, 
      from = "qpcr_bd_results") %>%
  collect()

ves_table <-
  tbl(src = ribbitr_connection, 
      from = "ves") %>%
  collect()
```

## Munge data tables to get data we need. 
```{r}
munged_location_table <-
  location_table 

munged_region_table <-
  region_table %>%
    filter(region == 'pennsylvania') %>%
  right_join(munged_location_table, by = 'location_id')

munged_site_table <-
  site_table %>%
    right_join(., munged_region_table, 
               by = 'region_id')

munged_visit_table <-
  visit_table %>%
    right_join(., munged_site_table, 
               by = 'site_id') %>%
  filter(year(date) > 2021) # this gets only data from 2023 and 2024
  
munged_survey_table <-
  survey_table %>% 
    right_join(., munged_visit_table, 
               by = 'visit_id')

munged_capture_table <-
  capture_table %>%
    right_join(., munged_survey_table, 
               by = 'survey_id') %>% 
# this line of code is here as it appears that there is a single recore of rana muscosa (only lives in sierra nevadas) that has been replicated throughout the database as from Pennsylvania.  super strange, but not gonna deal with this now...   
    drop_na(time_of_capture)

munged_qpcr_bd_results_table <-
  qpcr_bd_results_table %>%
    right_join(., munged_capture_table, 
               by = "bd_swab_id") %>%
    drop_na(bd_swab_id) %>% 
    filter(!str_detect(bd_swab_id, "-LA"))


munged_ves_table <-  
  ves_table %>% 
    left_join(munged_survey_table, by = 'survey_id') %>% 
    select(species_ves, 
           count, 
           date, 
           site, 
           observer = observer.x) %>%
    drop_na(site)
```

## Get table to send to cori and miranda. 
```{r}
#munged_qpcr_bd_results_table %>%
  select(bd_swab_id:comments, 
         site, 
         region,
         location) %>% view()
  write_csv("output/RIBBiTR_bd_swabs_from_2023_and_2024.csv")
```

## filter capture data to only get PA samples and anything from spring 2023 on. 
```{r}
munged_capture_table %>%
  filter(region == 'pennsylvania') %>%
  select(date, everything()) %>% 
  select(date, 
         species = species_capture, 
         site)
```

## Read in sample collection data from 2024 thus far (3 surveys.)
```{r}
processing_data_2024 <-
  read_csv('data/a8b92d20-97fc-4920-99e3-16379506f3a8/6_penn_sampleprocessing_amphibian_capture_survey_collections.csv')

pre_processing_data_2024 <-
  read_csv('data/a8b92d20-97fc-4920-99e3-16379506f3a8/6_penn_sampleprocessing.csv')

data_2024 <-
  left_join(processing_data_2024, pre_processing_data_2024, by = "_record_id") %>%
  select(date, 
         species = species_capture, 
         site = location)
```

```{r}
data_2024 %>%
  rbind(munged_capture_table %>%
  filter(region == 'pennsylvania') %>%
  select(date, everything()) %>% 
  select(date, 
         species = species_capture, 
         site)) %>%
  mutate(species = tolower(species)) %>%
  mutate(site = tolower(site)) %>%
  mutate(species = str_replace(species, "_", " ")) %>%
  mutate(site = str_replace(site, "_", " ")) %>%
  group_by(date, species, site) %>%
  tally() %>%
  mutate(date = as.Date(date, '%Y-%m-%d')) %>% 
  filter(species != 'negative control')  %>%
  arrange(site, date) %>% 
  ungroup() %>%
  pivot_wider(names_from = 'species', 
              values_from = 'n') %>% 
  mutate(month = month(date)) %>% 
  mutate(year = year(date)) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(season = if_else(month < 6, "spring", 
                          if_else(month < 9, "summer", "fall"))) %>% 
  select(-date, -month) %>%
  group_by(year, season, site) %>%
  summarise_all(sum) %>% 
  write_csv('output/sampling-species-per-time-period-spring2023-present.csv')
```

## Determine the number of individiuals we captured at each site in each year and plot it. 
```{r}
data_2024 %>%
  rbind(munged_capture_table %>%
  filter(region == 'pennsylvania') %>%
  select(date, everything()) %>% 
  select(date, 
         species = species_capture, 
         site)) %>%
  mutate(species = tolower(species)) %>%
  mutate(site = tolower(site)) %>%
  mutate(species = str_replace(species, "_", " ")) %>%
  mutate(site = str_replace(site, "_", " ")) %>%
  group_by(date, species, site) %>%
  tally() %>%
  mutate(date = as.Date(date, '%Y-%m-%d')) %>% 
  filter(species != 'negative control')  %>%
  arrange(site, date) %>% 
  ungroup() %>%
  pivot_wider(names_from = 'species', 
              values_from = 'n') %>% 
  mutate(month = month(date)) %>% 
  mutate(year = year(date)) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(season = if_else(month < 6, "spring", 
                          if_else(month < 9, "summer", "fall"))) %>% 
  select(-date) %>%
  group_by(year, season, site) %>%
  summarise_all(sum) %>%
  clean_names() %>%
  pivot_longer(cols = c(pseudacris_crucifer:eurycea_bislineata),
               names_to = 'species', 
               values_to = 'count') %>% 
  ungroup() %>%
  filter(species != 'unknown_species') %>%
  mutate(species = str_replace(species, "bislan", "bislin")) %>%
  filter(species != 'rana_catesbeiana_x_rana_clamitans_possibly') %>% 
  select(-site, -month) %>% 
  group_by(year, season, species) %>%
  summarise_all(sum) %>%
  mutate(season = factor(season, levels = c("spring", "summer", "fall"))) %>%
  mutate(spec.type = if_else(species == 'rana_catesbeiana' |
                               species == 'rana_pipiens' |
                               species == 'rana_clamitans' |
                               species == 'pseudacris_crucifer', "focal", "non-focal")) %>%
  filter(species != 'rana_sp') %>%
  ggplot() +
  geom_bar(aes(x = season, 
                y = count, 
               fill = if_else(spec.type == 'focal', species, 'non_focal')),
           colour = 'black',
           stat = 'identity') +
  scale_fill_manual(values = c("lightgrey", "#56B4E9", "#009E73", "#CC79A7", "#D55E00")) +
  facet_wrap(~year) +
  labs(fill = "Captured Amphbian Species") +
  theme_bw()



saveWidget(ggplotly(), file = "output/interactive_capture_plot_by_year_Penn.html")

ggsave('output/interactive_capture_plot_by_year_Penn.tiff')
```

## Determine the number of each sample type that was collected from PLE
```{r}
munged_capture_table %>% 
  select(species_capture, date, bd_swab_id:antibody_id_4) %>%
  drop_na(bd_swab_id) %>% 
  mutate(year = year(date)) %>% 
  distinct() %>% 
  mutate(across(contains("_id"), ~ if_else(is.na(.), 0, 1))) %>% # this little piece of code changes all na to 0 and all filled variables to 1 so that we can add them up later!
  pivot_longer(cols = bd_swab_id:antibody_id_4, 
               values_to = 'count') %>%
  select(-date) %>% 
  mutate(name = str_replace(name, "_[0-9]", "")) %>%
  group_by(species_capture, year, name) %>%
  summarise_all(sum) %>%
  ungroup() %>%
  select(species = species_capture, 
         everything()) %>%
  filter(count > 0) %>% 
  filter(species != 'unknown_species') %>%
  mutate(species = str_replace(species, "bislan", "bislin")) %>%
  filter(species != 'Rana_catesbeiana_x_Rana_clamitans_(possibly)') %>%
  filter(species != 'Rana_sp') %>%
  filter(species != 'Desmognathaus_sp.') %>%
  mutate(spec.type = if_else(species == 'rana_catesbeiana' |
                               species == 'rana_pipiens' |
                               species == 'rana_clamitans' |
                               species == 'pseudacris_crucifer', "focal", "non-focal")) %>%
  filter(name != 'crispr_id') %>%
  ggplot() +
  geom_bar(aes(x = name, 
               y = count, 
               fill = if_else(spec.type == 'focal', species, 'non_focal')),
           color = 'black',
           stat = 'identity') +
    facet_wrap(~year, 
               ncol = 1) +
  scale_fill_manual(values = c("lightgrey", "#56B4E9", "#009E73", "#CC79A7", "#D55E00")) +
  labs(fill = "Processed Amphbian Species",
       x = "Sample Type") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   vjust = 1))
  

saveWidget(ggplotly(), file = "output/interactive_process_plot_by_year_Penn.html")

ggsave('output/interactive_process_plot_by_year_Penn.tiff')
```

## make VES figure. 
```{r}
munged_ves_table %>%
  select(species = species_ves, 
         everything()) %>%
  group_by(date, species, site, observer) %>%
  summarise_all(mean) %>%
  mutate(date = as.Date(date, '%Y-%m-%d')) %>% 
  arrange(site, date) %>% 
  ungroup() %>%
  pivot_wider(names_from = 'species', 
              values_from = 'count') %>% 
  mutate(month = month(date)) %>% 
  mutate(year = year(date)) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(season = if_else(month < 6, "spring", 
                          if_else(month < 9, "summer", "fall"))) %>% 
  select(-date, -observer, -month) %>%
  group_by(year, season, site) %>%
  summarise_all(sum) %>%
  clean_names() %>% 
  pivot_longer(cols = c(rana_catesbeiana:pseudacris_crucifer),
               names_to = 'species', 
               values_to = 'count') %>% 
  ungroup() %>% 
  filter(species != 'unknown') %>%
  select(-site) %>% 
  group_by(year, season, species) %>%
  summarise_all(sum) %>%
  mutate(season = factor(season, levels = c("spring", "summer", "fall"))) %>%
  mutate(spec.type = if_else(species == 'rana_catesbeiana' |
                               species == 'rana_pipiens' |
                               species == 'rana_clamitans' |
                               species == 'pseudacris_crucifer', "focal", "non-focal")) %>%
  filter(species != 'rana_sp') %>%
  ggplot() +
  geom_bar(aes(x = season, 
                y = count, 
               fill = if_else(spec.type == 'focal', species, 'non_focal')),
           colour = 'black',
           stat = 'identity') +
  scale_fill_manual(values = c("lightgrey", "#56B4E9", "#009E73", "#CC79A7", "#D55E00")) +
  facet_wrap(~year) +
  labs(fill = "Visually-encountered Amphbian Species") +
  theme_bw()

saveWidget(ggplotly(), file = "output/interactive_visual_plot_by_year_Penn.html")

ggsave('output/interactive_visual_plot_by_year_Penn.tiff')
```

## Make sample sheet for Patrick Kearns (29 April 2024; see email)
```{r}
munged_capture_table %>%
  select(date, everything()) %>%
  arrange(date) %>%
  filter(year(date) > 2022) %>%
  select(date, bd_swab_id, species_capture, microbiome_swab_id, mucosome_id, bacterial_swab_id) %>%
  drop_na(bd_swab_id) %>% 
  filter(!is.na(microbiome_swab_id) |
           !is.na(mucosome_id) |
           !is.na(bacterial_swab_id))  %>% 
  write_csv('output/sample-sheet-for-umass-kearns-mcdonnell-29apr24.csv')
```

